```{r, include = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(spatstat)
library(choroplethrMaps)
library(choroplethr)
library(devtools)
#devtools::install_github("amd112/rseiAnalysis")
#library("rseiAnalysis")
```

# Approaches and Methods {#ref-labels}

## Level of Analysis
 
The questions we would like to pose - how the distributions of toxicity that individuals experience over time are predicted by their complex, multidimensional identities - is inherently a person level analysis. 
 
~~insert more explanation of level of analysis choices. 
  
This analysis depends on two data sources, the disaggregated RSEI toxic release data as compiled to contain only releases that are consistently reported between 1990 and 2010, as well as relevant demographic information from the census. 
 
RSEI environmental data can be obtained at extremely fine level (the 800 meter grid across the United States,) but the finest grain Census data is available at is the block level, which contains between 0 and a few hundred people. At such low geographic levels, few variables are available for demographics due to identifiability concerns. Additionally, few cross tabulations are available, due to concerns of identifiability. Using a low level of geography (like census blocks or block groups) is important for the environmental aspect of this analysis, since environmental hazards can be very localized, especially along neighborhood lines in urban areas. 
 
Unfortunately, the availability of cross tabulations is equally important to the goal of this work in examining inequality of environmental burden held by minorty groups in America. The intersection of social identies, especially those steeped in systems of oppression, is extremely important for identifying unequal burdens. For example, low income populations across the board may be more likely to experience environmental hazards, but low income minority populations could be much more likely than low income white populations to experience extreme hazard. The intersections of demographic charictersitics, such as race and income or race and education are likely to be important in teasing out differences in toxicity for a given person or region. 
 
~~insert the look at how the distributions differ when aggregated at each census level.


##Need to get census data for income, education, and cross tabs on race and income. 
##Need to figure out how to do a time series analysis. do I want to do a 

## I want to do a non-parametric simulation of what would have happened to the distributions if they had stayed in the same place as they were in the original distribution. 
How does the comparison there work? Do I just say wow look the real new distribution is actually significantly better, showing that there is real improvement for people of color. If we break it up by the states that implemented laws regarding environmental justice do we see a significant difference in the difference in differences? 

AKAA
Step 1) simulate the new distribution and see how the new and old dists. compare to the simulated new dist. 
Step 2) do we see that hte new is better than the simulated new?
Step 3) do we see the same "positional advancement" in the states with and without laws regarding ej?
Step 4) can we test that there is a difference in "positional advancement" (aka difference in simulated and real means)
Step 5) can we test that there is a difference in "positional advancement" for the states that have good general environmental laws?
Conclusions: there has been an improvement in the position of poc in the american toxicity distribution, and it is not/may be related to the specific ej laws.

```{r, include = FALSE}
t1990 <- read_csv("data/toxic_1990_1990_blockgroup.csv")
t1990$block = substr(t1990$block, 2, 12)
c1990 <- read_csv("data/LTDB_Std_1990_fullcount.csv")
t1990 = merge(t1990, c1990, by.x = "block", by.y = "TRTID10")
t1990$lconcentration = log(t1990$concentration)
t1990 = select(t1990, block, lconcentration, concentration, POP90, NHWHT90, NHBLK90, HISP90, OWN90, RENT90)
names(t1990) = c("block", "lconcentration", "concentration", "tot_pop", "white", "black", "hispanic", "owners", "renters")
t2000 <- read_csv("data/toxic_2000_2000_blockgroup.csv")
#t2000 = aggregateGeography(t2000, "tract")
c2000 <- read_csv("data/LTDB_Std_2000_fullcount.csv")
t2000 = merge(t2000, c2000, by.x = "block", by.y = "TRTID10")
t2000$lconcentration = log(t2000$concentration)
t2000 = select(t2000, block, lconcentration, concentration, POP00, NHWHT00, NHBLK00, HISP00, OWN00, RENT00)
names(t2000) = c("block", "lconcentration", "concentration", "tot_pop", "white", "black", "hispanic", "owners", "renters")
t2010 <- read_csv("data/toxic_2010_2010_blockgroup.csv")
#t2010 = aggregateGeography(t2010, "tract")
c2010 <- read_csv("data/LTDB_Std_2010_fullcount.csv", col_types = cols(tractid = col_character()))
c2010$tractid = str_pad(c2010$tractid, 11, pad = "0")
t2010 = merge(t2010, c2010, by.x = "block", by.y = "tractid")
t2010$lconcentration = log(t2010$concentration)
t2010 = select(t2010, block, lconcentration, concentration, pop10, nhwht10, nhblk10, hisp10, own10, rent10)
names(t2010) = c("block", "lconcentration", "concentration", "tot_pop", "white", "black", "hispanic", "owners", "renters")
```