```{r, include = FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(magrittr)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(spatstat)
library(choroplethrMaps)
library(choroplethr)
library(tidyr)
library(devtools)
devtools::install_github("amd112/rseiAnalysis")
library("rseiAnalysis")
#needed for RSEI
library(plyr)
```

# Approaches and Methods {#ref-labels}

## Geographic Level of Analysis

The questions we would like to pose - how the distributions of toxicity that individuals experience over time are predicted by their complex, multidimensional identities - is inherently intended to be a person level analysis. That intent may not be achievable given the available data. 

This analysis depends on two data sources, the disaggregated RSEI toxic release data as compiled to contain only releases that are consistently reported between 1990 and 2010, as well as relevant demographic information from the census. 

RSEI environmental data can be obtained at extremely fine level (the 800 meter grid across the United States,) but the finest grain Census data is available at is the block level, which contains between 0 and a few hundred people. At such low geographic levels, few variables are available for demographics due to identifiability concerns. Additionally, few cross tabulations are available, due to concerns of identifiability. Using a low level of geography (like census blocks or block groups) is important for the environmental aspect of this analysis, since environmental hazards can be very localized, especially along neighborhood lines in urban areas. 

Unfortunately, the availability of cross tabulations is equally important to the goal of this work in examining inequality of environmental burden held by minority groups in America. The intersection of social identies, especially those steeped in systems of oppression, is extremely important for identifying unequal burdens. For example, low income populations across the board may be more likely to experience environmental hazards, but low income minority populations could be much more likely than low income white populations to experience extreme hazard. The intersections of demographic charictersitics, such as race and income or race and education are likely to be important in teasing out differences true inequality burden. 

To combine the data, we assign the aggregated releases to their respective geographies. For each census geography, we also have estimates of number of various population groups. From there, we move to a person level analysis by assigning each of the people the toxicity for the geography they originated in. By assigning each person the toxicity of their block group, we can aggregate nationally to find the distribution of toxicity that each group experiences. This approach is restricted in ability to approach the problem in an intersectional manner, since we can only build a distribution for each of the crosstabs we have available. For higher levels of geography (where we might, for example, have race by income) we would be able to build national distributions for each income by race group.

In choosing the level of aggregation at which to assign toxicity, in order to balance the needs of accuracy of toxicity and availability of cross tabulations, we create the overall toxicity distribution for Americans by assigning each individual a toxicity level for each level of geography. Using block group as the smallest form of geography, and aggregating up to state (including tract and county in between) we see how the distribution changes at each level of aggregation. 

```{r, echo = FALSE, fig.width=6, fig.height=3, message = FALSE}
#import toxicity data
bg2010 = read_csv("data/toxic_2010_2010_blockgroup.csv")

#import census data
census_bg = read_csv("data/ecologicalFallacy/census_bg_race.csv", skip = 1)
census_tract = read_csv("data/ecologicalFallacy/census_tract_race.csv", skip = 1)
census_county = read_csv("data/ecologicalFallacy/census_county_race.csv", skip = 1)
census_state = read_csv("data/ecologicalFallacy/census_state_race.csv", skip = 1)

#munge census data
census_bg = select(census_bg, Geo_FIPS, SE_T013_001, SE_T013_002, SE_T013_003)
names(census_bg)  = c("fips", "total", "white", "black")
census_tract = select(census_tract, Geo_FIPS, SE_T013_001, SE_T013_002, SE_T013_003)
names(census_tract)  = c("fips", "total", "white", "black")
census_county = select(census_county, Geo_FIPS, SE_T013_001, SE_T013_002, SE_T013_003)
names(census_county)  = c("fips", "total", "white", "black")
census_state = select(census_state, Geo_FIPS, SE_T013_001, SE_T013_002, SE_T013_003)
names(census_state)  = c("fips", "total", "white", "black")

#merge census and toxicity data
tract2010 = aggregateGeography(bg2010, "tract")
tract2010 = merge(tract2010, census_tract, by.x = "block", by.y = "fips")
county2010 = aggregateGeography(tract2010, "county")
county2010 = merge(county2010, census_county, by.x = "block", by.y = "fips")
state2010 = aggregateGeography(county2010, "state")
state2010 = merge(state2010, census_state, by.x = "block", by.y = "fips")
bg2010 = merge(bg2010, census_bg, by.x = "block", by.y = "fips")

#plot national distributions
g = ggplot() + geom_density(aes(log(county2010$concentration), weight = county2010$total/sum(county2010$total), color = "County"), bw = 0.2)
g = g + geom_density(aes(log(state2010$concentration), weight = state2010$total/sum(state2010$total), color = "State"), bw = 0.35)
g = g + geom_density(aes(log(bg2010$concentration), weight = bg2010$total/sum(bg2010$total), color = "Block Group"), bw = 0.05)
g = g + geom_density(aes(log(tract2010$concentration), weight = tract2010$total/sum(tract2010$total), color = "Tract"), bw = 0.05)
g = g + labs(x = "Log Toxicity", colour = "Geographic Level", caption = "(as assigned by different levels of geography)", title = "Distribution of Toxicity for US Population")
g = g + theme(axis.text=element_text(size = 7), legend.text = element_text(size = 7), legend.title = element_text(size = 8))
g
```

We see expected results when looking at the aggregation at state level. Given that we are assigning each individual the mean toxicity in their entire state, we are eliminating most of the variation from the data. States seem to have overall similar toxicity levels, but interestingly the tract data seems to build a distribution very similar to the block group level assignment. This may be because the block group level is aggregating a large group of our fine grain toxicity data, and has therefore already lost the street block by street block variation that we had deemed so crucial, so aggregating several block groups leaves us in a similar neighborhood level of aggregation. 

## Temporal Level of Analysis 

## Non-Parametric Analysis

## Trends Over Time

```{r, include = FALSE}
t1990 <- read_csv("data/toxic_1990_1990_blockgroup.csv")
t1990$block = substr(t1990$block, 2, 12)
c1990 <- read_csv("data/LTDB_Std_1990_fullcount.csv")
t1990 = merge(t1990, c1990, by.x = "block", by.y = "TRTID10")
t1990$lconcentration = log(t1990$concentration)
t1990 = select(t1990, block, lconcentration, concentration, POP90, NHWHT90, NHBLK90, HISP90, OWN90, RENT90, area)
names(t1990) = c("block", "lconcentration", "concentration", "tot_pop", "white", "black", "hispanic", "owners", "renters", "area")

t2000 <- read_csv("data/toxic_2000_2000_blockgroup.csv")
t2000 = aggregateGeography(t2000, "tract")
c2000 <- read_csv("data/LTDB_Std_2000_fullcount.csv")
t2000 = merge(t2000, c2000, by.x = "block", by.y = "TRTID10")
t2000$lconcentration = log(t2000$concentration)
t2000 = select(t2000, block, lconcentration, concentration, POP00, NHWHT00, NHBLK00, HISP00, OWN00, RENT00, area)
names(t2000) = c("block", "lconcentration", "concentration", "tot_pop", "white", "black", "hispanic", "owners", "renters", "area")

t2010 <- read_csv("data/toxic_2010_2010_blockgroup.csv")
t2010 = aggregateGeography(t2010, "tract")
c2010 <- read_csv("data/LTDB_Std_2010_fullcount.csv", col_types = cols(tractid = col_character()))
c2010$tractid = str_pad(c2010$tractid, 11, pad = "0")
t2010 = merge(t2010, c2010, by.x = "block", by.y = "tractid")
t2010$lconcentration = log(t2010$concentration)
t2010 = select(t2010, block, lconcentration, concentration, pop10, nhwht10, nhblk10, hisp10, own10, rent10, area)
names(t2010) = c("block", "lconcentration", "concentration", "tot_pop", "white", "black", "hispanic", "owners", "renters", "area")
```

```{r}
vars = c("white", "black", "hispanic", "owners", "renters")
p1990 = t1990
p1990[vars] = p1990[vars]/p1990$tot_pop
p1990$density = p1990$tot_pop/p1990$area
m = lm(lconcentration ~ 1 + owners + renters + black + hispanic + density, data = p1990)

p2010 = t2010
p2010[vars] = p2010[vars]/p2010$tot_pop
p2010$density = p2010$tot_pop/p2010$area
m = lm(lconcentration ~ 1 + owners + renters + black + hispanic + density, data = p2010)
```
