---
author: 'Anne Driscoll'
date: 'May 2018'
institution: 'Duke University'
division: 'Trinity College of Arts and Sciences'
advisor: 'David Banks'
#altadvisor: 'Your Other Advisor'
# Delete line 7 if you only have one advisor
committeememberone: 'Committeemember O. Name'
committeemembertwo: 'Committeemember T. Name'
dus: 'Mine Cetinkaya Rundel'
department: 'Department of Statistical Science'
degree: 'Bachelor of Science in Statistical Science'
title: 'Examining Injustice in Environmental Toxicity'
knit: "bookdown::render_book"
site: bookdown::bookdown_site
output: 
#  thesisdowndss::thesis_pdf: default
  thesisdowndss::thesis_gitbook: default
#  thesisdowndss::thesis_word: default
#  thesisdowndss::thesis_epub: default
# If you are creating a PDF you'll need to write your preliminary content here or
# use code similar to line 20 for the files.  If you are producing in a different
# format than PDF, you can delete or ignore lines 20-31 in this YAML header.
abstract: |
  `r if(knitr:::is_latex_output()) paste(readLines("00-abstract.Rmd"), collapse = '\n  ')`
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab is 
# needed on the line after the |.
acknowledgements: |
  I want to thank a few people.
dedication: |
  You can have a dedication here if you wish. 
preface: |
  This is an example of a thesis setup to use the reed thesis document class
  (for LaTeX) and the R bookdown package, in general.
bibliography: bib/thesis.bib
# Refer to your specific bibliography file in the line above.
csl: csl/apa.csl
# Download your specific csl file and refer to it in the line above.
lot: true
lof: true
space_between_paragraphs: true
# Delete the # at the beginning of the previous line if you'd like
# to have a blank new line between each paragraph
#header-includes:
#- \usepackage{tikz}
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete lines 17 and 18 above or add a # before them to comment them out.  If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.
-->

<!--
If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r libraries, include_packages, include = FALSE}
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(thesisdowndss))
  devtools::install_github("mine-cetinkaya-rundel/thesisdowndss")
library(thesisdowndss)
library(knitr)
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
knitr::opts_chunk$set(echo = FALSE) 
library(magrittr)
library(readr)
library(ggplot2)
library(dplyr)
library(plyr)
library(stringr)
library(spatstat)
library(readr)
library(tidyr)
library(devtools)
devtools::install_github("amd112/rseiAnalysis")
library("rseiAnalysis")
#needed for RSEI
set.seed(4012)
library(data.table)
library(grid)
library(gridExtra)
```

```{r readInData, include = FALSE}
t1990 = as.data.frame(fread("data/toxic/tract/toxic_1990_2010_tract.csv", data.table = FALSE))
t1991 = as.data.frame(fread("data/toxic/tract/toxic_1991_2010_tract.csv", data.table = FALSE))
t1992 = as.data.frame(fread("data/toxic/tract/toxic_1992_2010_tract.csv", data.table = FALSE))
t1993 = as.data.frame(fread("data/toxic/tract/toxic_1993_2010_tract.csv", data.table = FALSE))
t1994 = as.data.frame(fread("data/toxic/tract/toxic_1994_2010_tract.csv", data.table = FALSE))
#t1995 = as.data.frame(fread("data/toxic/tract/toxic_1995_2010_tract.csv", data.table = FALSE))
t1996 = as.data.frame(fread("data/toxic/tract/toxic_1996_2010_tract.csv", data.table = FALSE))
t1997 = as.data.frame(fread("data/toxic/tract/toxic_1997_2010_tract.csv", data.table = FALSE))
t1998 = as.data.frame(fread("data/toxic/tract/toxic_1998_2010_tract.csv", data.table = FALSE))
t1999 = as.data.frame(fread("data/toxic/tract/toxic_1999_2010_tract.csv", data.table = FALSE))
t2000 = as.data.frame(fread("data/toxic/tract/toxic_2000_2010_tract.csv", data.table = FALSE))
t2001 = as.data.frame(fread("data/toxic/tract/toxic_2001_2010_tract.csv", data.table = FALSE))
t2002 = as.data.frame(fread("data/toxic/tract/toxic_2002_2010_tract.csv", data.table = FALSE))
t2003 = as.data.frame(fread("data/toxic/tract/toxic_2003_2010_tract.csv", data.table = FALSE))
#t2004 = as.data.frame(fread("data/toxic/tract/toxic_2004_2010_tract.csv", data.table = FALSE))
t2005 = as.data.frame(fread("data/toxic/tract/toxic_2005_2010_tract.csv", data.table = FALSE))
t2006 = as.data.frame(fread("data/toxic/tract/toxic_2006_2010_tract.csv", data.table = FALSE))
t2007 = as.data.frame(fread("data/toxic/tract/toxic_2007_2010_tract.csv", data.table = FALSE))
t2008 = as.data.frame(fread("data/toxic/tract/toxic_2008_2010_tract.csv", data.table = FALSE))
t2009 = as.data.frame(fread("data/toxic/tract/toxic_2009_2010_tract.csv", data.table = FALSE))
t2010 = as.data.frame(fread("data/toxic/tract/toxic_2010_2010_tract.csv", data.table = FALSE))
t2011 = as.data.frame(fread("data/toxic/tract/toxic_2011_2010_tract.csv", data.table = FALSE))
t2012 = as.data.frame(fread("data/toxic/tract/toxic_2012_2010_tract.csv", data.table = FALSE))
t2013 = as.data.frame(fread("data/toxic/tract/toxic_2013_2010_tract.csv", data.table = FALSE))
t2014 = as.data.frame(fread("data/toxic/tract/toxic_2014_2010_tract.csv", data.table = FALSE))

read_in = function(year, tox) {
  data = as.data.frame(fread(paste0("data/census/race/race_tract_", year, ".csv"), data.table = FALSE))
  data = data[complete.cases(data), ]
  data$id = str_pad(data$id, 11, pad = "0", side = "left")
  data = merge(data, tox, by.x = "id", by.y = "block")
  data$lconcentration = log(data$concentration)
  data$year = year
  return(data)
}

#r1990 = read_in(1990, t1990)
#r1991 = read_in(1991, t1991)
#r1992 = read_in(1992, t1992)
#r1993 = read_in(1993, t1993)
#r1994 = read_in(1994, t1994)
#r1996 = read_in(1996, t1996)
#r1997 = read_in(1997, t1997)
#r1998 = read_in(1998, t1998)
#r1999 = read_in(1999, t1999)
#r2000 = read_in(2000, t2000)
#r2001 = read_in(2001, t2001)
#r2002 = read_in(2002, t2002)
#r2003 = read_in(2003, t2003)
#r2005 = read_in(2005, t2005)
#r2006 = read_in(2006, t2006)
#r2007 = read_in(2007, t2007)
#r2008 = read_in(2008, t2008)
#r2009 = read_in(2009, t2009)
#r2010 = read_in(2010, t2010)

#race_data = rbind.fill(r1990, r1991, r1992, r1993, r1994, r1996, r1997, r1998, r1999, r2000, r2001, r2002, r2003, r2005, r2006, r2007, r2008, r2009, r2010)

read_in = function(year, tox) {
  data = as.data.frame(fread(paste0("data/census/hisp_pov/race_pov_tract_", year, ".csv"), data.table = FALSE))
  data$hisp = 
  data = data[complete.cases(data), ]
  data$id = str_pad(data$id, 11, pad = "0", side = "left")
  data = merge(data, tox, by.x = "id", by.y = "block")
  data$year = year
  return(data)
}

#p1990 = read_in(1990, t1990)
#p1991 = read_in(1991, t1991)
#p1992 = read_in(1992, t1992)
#p1993 = read_in(1993, t1993)
#p1994 = read_in(1994, t1994)
#p1996 = read_in(1996, t1996)
#p1997 = read_in(1997, t1997)
#p1998 = read_in(1998, t1998)
#p1999 = read_in(1999, t1999)
#p2000 = read_in(2000, t2000)
#p2001 = read_in(2001, t2001)
#p2002 = read_in(2002, t2002)
#p2003 = read_in(2003, t2003)
#p2005 = read_in(2005, t2005)
#p2006 = read_in(2006, t2006)
#p2007 = read_in(2007, t2007)
#p2008 = read_in(2008, t2008)
#p2009 = read_in(2009, t2009)
#p2010 = read_in(2010, t2010)

#pov_data = rbind.fill(p1990, p1991, p1992, p1993, p1994, p1996, p1997, p1998, p1999, p2000, p2001, p2002, p2003, p2005, p2006, p2007, p2008, p2009, p2010)

#Environmental justice literature has focused largely on the motivations behind facility siting that lead to inequality and policy approaches to help fix the disproportionate burden of environmental hazards on minority communities. The questions being asked are critical to creating policy that helps correct the problem of hazard allocation. A gap in the literature is the need for an investigation of the dynamic world of hazard allocation over time. 

# what is natural sorting, How are env dumps created?, Breaking the self fulfilling cycle

read_in = function(year, tox) {
  data = as.data.frame(fread(paste0("data/census/hisp/hisp_tract_", year, ".csv"), data.table = FALSE))
  data = data[complete.cases(data), ]
  data$other_sum = rowSums(data[, 5:7])
  data = select(data, id, pop, white, black, hispanic, other_sum)
  names(data) = c("id", "pop", "white", "black", "hispanic", "other")
  data$id = str_pad(data$id, 11, pad = "0", side = "left")
  data = merge(data, tox, by.x = "id", by.y = "block")
  data$year = year
  return(data)
}

h1990 = read_in(1990, t1990)
h1991 = read_in(1991, t1991)
h1992 = read_in(1992, t1992)
h1993 = read_in(1993, t1993)
h1994 = read_in(1994, t1994)
h1996 = read_in(1996, t1996)
h1997 = read_in(1997, t1997)
h1998 = read_in(1998, t1998)
h1999 = read_in(1999, t1999)
h2000 = read_in(2000, t2000)
h2001 = read_in(2001, t2001)
h2002 = read_in(2002, t2002)
h2003 = read_in(2003, t2003)
h2005 = read_in(2005, t2005)
h2006 = read_in(2006, t2006)
h2007 = read_in(2007, t2007)
h2008 = read_in(2008, t2008)
h2009 = read_in(2009, t2009)
h2010 = read_in(2010, t2010)

hisp_data = rbind.fill(h1990, h1991, h1992, h1993, h1994, h1996, h1997, h1998, h1999, h2000, h2001, h2002, h2003, h2005, h2006, h2007, h2008, h2009, h2010)
```


<!-- You'll need to include the order that you'd like Rmd files to appear in the _bookdown.yml file for
PDF files and also delete the # before rmd_files: there.  You'll want to not include 00(two-hyphens)prelim.Rmd
and 00-abstract.Rmd since they are handled in the YAML above differently for the PDF version.
-->

<!-- The {.unnumbered} option here means that the introduction will be "Chapter 0." You can also use {-} for no numbers
on chapters.
-->

# Abstract {.unnumbered}

This thesis seeks to address the question of how the toxicity experienced by minority communities has changed. Specifically, we look for substantial reductions in relative environmental toxicity, when shifts are occur, and what traits that are predictive of high environmental toxicity.

### Keywords {.unnumbered}
Environmental Inequality, Racial Inequality, Non-parametric analysis
